The landscape of chloroplast genome assembly tools

### 2020-05-19 manuscript/reviewer_response1.tex

\documentclass[a4paper,10pt]{scrartcl}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\title{Response to Reviewers}
\author{Markus J. Ankenbrand and Frank F\"{o}rster for the authors}
\date{\today}
%
\begin{document}
\maketitle
\section{Editor}
We have now received reports from two referees, whose comments are available below.
As you will see from the reports, both referees raised serious concerns that the benchmark hasn't provided sufficient advance to be useful for the field.
At this stage, therefore, we are not persuaded that an offer of publication in Genome Biology would be justified.
Unless you can address all of the issues raised by the referees I am afraid that we would not be able to consider this work again.
In particular we ask that you provide additional analyses as requested by both referees to convince them of the advance.
The revised manuscript should address in full all the points raised by both referees; a separate list of the revisions made, specifying where in the manuscript the changes have been made, would also be helpful.
Please note that all data generated for the manuscript should be deposited in a relevant public repository, with accession codes listed in the manuscript's Availability of Data and Materials section.
All other supporting data should also be signposted.
If you decide to submit a revised version, you will need to  this as a new manuscript, although we would consider it in the same way as we do all revised manuscripts.
The exception is that we will take into account any work published in the intervening period.
For the time being, however, we hope that you will find our referees' comments helpful in deciding how to proceed.
Please feel free to contact us if you have any questions.

\section{Reviewer 1}
Methods are appropriate but same authors developed ChloroExtractor, an important tool used as a benchmark in this study.
Authors compare seven chloroplast genome assembly tools.
So, this reviewer is not sure of novelty of this manuscript.
So, authors haven't addressed the important question "Does the work represent a significant advance over previously published studies?".

\textbf{
Indeed the novelty in this article is not a new approach to the task of chloroplast genome assembly but rather a thourough benchmark of existing tools (including our tool ChloroExtractor).
This represents a substantial advance for the field as there are competing tools published for the same task without any information about their relative performances.
Currently, researchers have to do this evaluation of tools themselves to find the appropriate tool which leads to duplicated efforts as this is done by every group separately and without sharing results.
If this evaluation is neglected a suboptimal tool might be used which can lead to additional manual efforts in finishing or additional costs if more data is generated.
Thus, clear guidelines to the end user (and important directions for improvement to the developers) are novel, important and benficial for the research community.
}

User experience and success rate should be determined by users and documented through satisfaction surveys.
This is especially important when authors emphasize that program can be used using standard computer infrastructure.

\textbf{
We agree that this would be the best way to measure and quantify user experience.
However, this is not feasible with limited budget and time (TODO: is this excuse good enough?) while the main focus of the work was on quantitative performance measures, anyway.
We decided to include usability metrics that were obtainable in a defined and structured way using the JOSS reviewer guidelines.
}

There is >1,000 sequenced chloroplast genomes, with a large majority not fully assembled or annotated.
So, authors have plenty of opportunity to demonstrate use of these new assembly tools.

\textbf{
Thanks for pointing this out.
We now include as a proof of principle the application of these tools on more than 100 novel datasets.
We thus clearly demonstrate that the insights from this article can be used to produce further novel results that can lead to biological insight.
}

It is well known that intergenic spacer regions are not conserved and not even a single intergenic spacer region is conserved within chloroplasts genomes of the graminae or grass genomes.
Intergenic spacer region occupies >50\% of the genome.
So, how does reference genome help in genome assembly?

\textbf{
This is a very interesting question.
We don't think that this is within the scope of our publication as we compare de novo assembly tools.
TODO: some people use it, some programs use it, it seems to work often but actually your argument is justification why we restricted our analysis to de novo tools.
}

In order to have real life comparison, authors could have sequenced few chloroplast genomes and used that data to compare different assembly tools and time taken for each program, pros and cons.

\textbf{
We already tested the tools on real datasets downloaded from SRA which is equivalent to customly sequenced data.
But we now include additional 100+ novel data sets with no reference chloroplast in the database.
}

- Is the paper of broad interest to others in the field, or of outstanding interest to a broad audience of biologists?
This is of interest to evolutionary biologists to study phylogenetics or for chloroplast genetic engineering.

\section{Reviewer 2}
This study provides an systematically comparison of the chloroplast assembly tools at first and show significant differences between the tested assemblers in terms of generating whole chloroplast genome sequences and computational requirements.
Moreover, the authors created docker images for each tested tool, which are available for the scientific community and improved reproducibility, thus large scale screening for chloroplasts as hidden treasures within genomic sequencing data is feasible.
This topic of the review paper is knowledgeable and helpful for all the scientists who are working in the field of plant chloroplast assembly.
However, the overall manuscript is not well written and contains limited information about systematically comparison of the chloroplast assembly tools and docker images for assembled reproducibility.

\textbf{
Thanks for your thorough review and the pointers that helped us improve the overall readability of the manuscript.
In addition we received feedback and edits from a native English speaker to further improve readabilty.
TODO: how do we adress the "limited information" aspect. We could link to the github repo that contains all data and code we used including the dockerfiles.
}

Specific points: The number of lines on the edge does not correspond to the maintext.

The major concerns about this manuscript can be found below and major revision is suggested before publication in "Genome Biology".

\begin{itemize}
    \item Page1  part of "General introduction and motivation"
    \item Line 4: It should be quote recent article
    \item Line 11: It should be quote recent article
    \item Page2
    \item Line 16: It should be IRa and IRb
    \item Line 18: It should be quote recent article
    \item Line 24-27: Re-write sentence
    \item Part of "Approaches to extracting chloroplasts from whole genome data"
    \item Line 6: Delete the dot before the quote
    \item Lines 7-9: Re-write sentence
    \item Part of "Purpose and scope of this study"
    \item Line 2: It should be in italic type
    \item Line 2: It should be genome
    \item Line 3: Raw genomic data sets should be filter out low-quality reads by quality control
    \item Lines 4 and 5: Many command-line tools of de novo assemble for chloroplast genome have all of these features, e.g., MITObim.
    \item Part of "Time requirements"
    \item Line 1: It should be among
    \item Line 4: Different tools with more than 1 gigabyte data sets should be tested, and Figure1 can`t give valuable information as no significant difference for time-consuming. The same question for Part of "Memory and CPU Usage". Why these seven assemblers were chosen to test? Figure should be "Fig." and table as "Table".
    \item Part of "Qualitative"
    \item Line 3: NGS data for crop genomes can`t be used for GetOrganelle? In this part, the same data set should be used to test for every assemblers better.
    \item Part of "Quantitative"  
    \item Part of "Simulated data"
    \item Line 1: Why not show the result for different data sets with every assemblers in Fig.3 and Table 2? Fig.3 and table 2 show no regular pattern in scores except the assembler GetOrganelle obtaining perfect.
    \item Part of "Real data sets"
    \item Line 3: The performance of GetOrganelle was bad, as only 199 circular assemblies out of a total of 356 assemblies that resulted in an output, the percent is 55.9\%, however it achieved a median scores of 99.7. The number "99.7" should to be "89.13", as show in table 3? The SD value of GetOrganelle is too large than the other three tools but it achieved the highest scores.
    \item Part of "Consistency"
    \item Line 3: As shown in the Fig.6, only two tools achieved R2  not equal 1 by re-running assemblies and comparison of the scores of two assemblers, however, GetOrganelle defined as the only tool that succeed in obtaining similar score for all assemblers, this conclusion is not rigorous.
\end{itemize}
\end{document}

### 2020-05-19 manuscript/bmc_article.tex
Update to latest manuscript version

\usepackage[noabbrev]{cleveref}
\usepackage{siunitx}
\DeclareSIUnit \basepair{bp}
\DeclareSIUnit \byte{B}
\DeclareSIPrefix\mebi{Mi}{10}
\DeclareSIPrefix\gibi{Gi}{10}

@@

\usepackage{todonotes}
\setlength{\marginparwidth}{3cm}

\newcounter{todocounter}
\newcommand{\ff}[1]
{\stepcounter{todocounter}
 \todo[color=blue!40,author=For Frank]{\thetodocounter: #1}
 }
\newcommand{\ak}[1]
{\stepcounter{todocounter}
 \todo[color=green!40,author=Arthur]{\thetodocounter: #1}
 }

@@

@@

@@

\author[
    addressref={aff1,aff3},
    noteref={n1},
    email={jan.freudenthal@uni-wuerzburg.de}
    ]{\inits{JAF}\fnm{Jan A} \snm{Freudenthal}}
\author[
    addressref={aff1,aff2},
    noteref={n1},
    email={simon-pfaff@gmx.de}
    ]{\inits{SP}\fnm{Simon} \snm{Pfaff}}
\author[
   addressref={aff1,aff3},
   email={niklas.terhoeven@uni-wuerzubrg.de}
]{\inits{NT}\fnm{Niklas} \snm{Terhoeven}}
\author[
   addressref={aff1},
   email={arthur.korte@uni-wuerzubrg.de}
]{\inits{AK}\fnm{Arthur} \snm{Korte}}
\author[
   addressref={aff1,aff3,aff2 },
   noteref={n2},
   email={markus.ankenbrand@uni-wuerzubrg.de}
]{\inits{MJA}\fnm{Markus J} \snm{Ankenbrand}}
\author[
   addressref={aff1,aff2,aff4},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   noteref={n2},                        % id's of article notes, if any
   email={frank.foerster@uni-wuerzburg.de}   % email address
]{\inits{FF}\fnm{Frank} \snm{Förster}}

@@

\address[id=aff3]{%
  \orgname{AnaLife Data Science},
  \street{Friedrich-Bergius-Ring 15},
  \postcode{97076}
  \city{W\"{u}rzburg},
  \cny{Germany}
}
\address[id=aff2]{%                           % unique id
  \orgname{Department of Bioinformatics, University of W\"{u}rzburg}, % university, etc
  \street{Biozentrum, Am Hubland},                     %
  \postcode{97074},                                % post or zip code
  \city{W\"{u}rzburg},                              % city
  \cny{Germany}                                    % country
}

\address[id=aff4]{%
  \orgname{Fraunhofer IME-BR},
  \street{Winchester Str.\ 2},
  \postcode{97076}
  \city{Gie\ss{}en},
  \cny{Germany}
}

@@

\begin{artnotes}
%\note{Sample of title note}     % note to the article
\note[id=n1]{Equal contributor} % note, connected to author
\note[id=n2]{Corresponding author} % note, connected to author
\end{artnotes}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstractbox}
\begin{abstract} % abstract
%One or two sentences providing a basic introduction to the field, comprehensible to a scientist in any discipline.
Chloroplasts are photosynthetic organelles in plant cells and contain their own genomic information.
That genome can be utilized in different scientific fields like phylogenetics or biotechnology.
%Two to three sentences of more detailed background, comprehensible to scientists in related disciplines.
Thus, different assemblers have been developed specialized in chloroplast assemblies.
Those assemblers often use the output of whole genome sequencing experiments as input.
Such sequencing data usually contain the complete chloroplast genome information, even if the sequencing aims for the core genome.
%One sentence clearly stating the general problem being addressed by this particular study.
Different assembly tools have never been systematically compared. %to each other concerning success rate or computational requirements.
%One sentence summarising the main result (with the words “here we show” or their equivalent).
Here we present a benchmark of seven chloroplast assembly tools, capable to succeed in more than \SI{60}{\percent} of real data sets.
%Two or three sentences explaining what the main result reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.
Our results show significant differences between the tested assemblers in terms of generating whole chloroplast genome sequences and computational requirements.
Moreover, we suggest further development to improve user experience and success rate.
%One or two sentences to put the results into a more general context. 
In terms of reproducibility, we created docker images for each tested tool, which are available for the scientific community.
Following the presented guidelines, users are able to analyze and screen data sets for chloroplast genomes using only standard computer infrastructure.
%Two or three sentences to provide a broader perspective, readily comprehensible to a scientist in any discipline, may be included in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced by their inclusion. 
Thus large scale screening for chloroplasts as hidden treasures within genomic sequencing data is feasible.

@@

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Introduction}
\subsection*{General introduction and motivation}
Chloroplasts are essential organelles present in plant cells and the cells of some protists.
Chloroplasts enable the conversion of light energy into chemical energy via photosynthesis.
They harbor their own ribosomes and a circular DNA genome usually with a size between \SIrange{120}{160}{\kilo\basepair} \cite{palmer_1985}.
Because of this small size, the chloroplast genome has been an early target for sequencing.
The first chloroplast genome sequences were obtained as early as 1986 \cite{ohyama_chloroplast_1986,shinozaki_complete_1986}.
These early efforts elucidated the general genome organization and structure of the chloroplast DNA. 
Chloroplast genome content and structure are reviewed for example in \cite{wicke_evolution_2011,green_chloroplast_2011}.
Chloroplast genomes are widely used for evolutionary analyses \cite{martin_plastid_2010,xiao-ming_inferring_2017}, barcoding \cite{kress_use_2005,hollingsworth_dna_2009,de_vere_dna_2015}, and meta-barcoding \cite{bell_review_2016,deiner_environmental_2017}.
Interesting aspects of chloroplast genomes are their small size (\SIrange{120}{160}{\kilo\basepair},\cite{palmer_1985}), caused through endosymbiotic gene transfer \cite{martin_evolutionary_2002,timmis_endosymbiotic_2004} and the low number of \numrange{100}{120} genes that are still encoded on the chloroplast genome \cite{wicke_evolution_2011}.
Despite the overall high conservation of the genome sequence, there are striking differences in the gene content between different groups (e.g. the loss of the whole \genename{ndh} gene family in Droseraceae \cite{nevill_plastome-wide_2019}). Even more extreme evolutionary cases, where chloroplasts show a very low GC content and a modified genetic code are described \cite{su_novel_2019}.

These differences call for comparative genomic approaches.
Given the small size, it is much easier to decipher the complete chloroplast genome than the complete core genome.
For example the \plant{At} core genome is approximately \SI{125}{\mega\basepair} in length \cite{schmuths2004,cao2011} while the size of the \plant{At} chloroplast genome with \SI{154}{\kilo\basepair} is more than \SI{800}{\times} smaller \cite{sato1999}.

Even if only a single chloroplast is located inside a plant cell, several hundreds copies of the chloroplast genome exists in each cell \cite{kumar_2014,bendich_1987}.
Therefore, many genome sequencing projects contain chloroplast reads as by-product.
In some cases the chloroplast data is even considered contamination and experimental protocols for reducing their content have been developed \cite{lutz_isolation_2011}.
An alternative approach to improve the assembly of the core genome would be to first resolve the chloroplast genome and afterwards use this information to remove those reads that map to the chloroplast genome.

Structurally, two inverted repeats (IR\textsubscript{A} and IR\textsubscript{B}) of \SIrange{10}{76}{\kilo\basepair} divide the chloroplast genome into a large (LSC) and a small single copy (SSC) region \cite{palmer_1985}.
Those large inverted repeats complicate automated resolution with short read technologies\cite{Wang2018}.
Moreover, the existence of different chloroplasts within a single individual, and thus multiple different chloroplast genomes, have been described for different plants \cite{corriveau_1988,Chat2002,Scar2016}.
Although the origin and evolutionary importance of this phenomena ---called heteroplasmy--- are only poorly understood, the assembly of whole chloroplast genomes might be hindered.

Databases exist containing short read data for species where no reference chloroplast sequence is publicly available, eg.\ the Sequence Read Archive at NCBI \cite{sra2010}.
The availability of whole chloroplast genomes would enable large scale comparative studies \cite{tonti-filippini_what_2017}.
Additionally, reconstructed full chloroplast genomes have been used as super-barcodes \cite{coissac_barcodes_2016}, for biotechnology applications and genetic engineering \cite{daniell_chloroplast_2016}.

\subsection*{Approaches to extracting chloroplasts from whole genome data}

Different strategies have been developed to assemble chloroplast genomes \cite{twyford_strategies_2017}.
In general, obtaining a chloroplast genome from WGS data requires two steps. First, the chloroplast reads have to be extracted from the mixed sequencing data. 
The second step is the assembly and resolution of the special circular structure including the inverted repeats.
The extraction of the reads can be achieved by mapping the reads to a reference chloroplast. \cite{Vinga2012}.
A different approach that does not perform alignments, relies on the higher coverage of chloroplast data in the whole genome sequencing data set\cite{Chan2013}.
Here, a $k$-mer analysis can be used to extract the most frequent reads.
An example for this is implemented in \ce{} \cite{j_ankenbrand_chloroextractor:_2018}. %\todo{ce ist hier der einzig explizit genannte Assembler, das wirkt wie bias :) entfernen? Nope... Steht klar ein Beispiel und wir haben nun mal den angesehen... Würde ich jetzt einmal so drin lassen. Alternativ nehmen wir noch einen dazu}
A third method combines both approaches by using a reference chloroplast as seed and simultaneously assembling the reads based on $k$-mers \cite{dierckxsens_novoplasty:_2017}.

@@

\subsection*{Purpose and scope of this study}
The goal of this study is to compare the effectiveness and efficiency of existing open source command-line tools to de-novo assemble whole chloroplast genomes from raw genomic data sets with minimal configuration. 
This includes no need for extensive data preparation, no need for a specific reference (apart from \plant{At}), no need to change default parameters, no manual finishing.
We further restricted our benchmark to paired end Illumina data sets as these are routinely generated by modern sequencing platforms \cite{Goodwin2016}.

In our opinion this reflects the most common use cases: (1) a user trying a tool quickly without digging into options for fine tuning and (2) large scale automatic applications.
Still, we acknowledge that the performance of the tools might be significantly improved by optimizing parameters (and references if applicable) for each data set specifically.
However, an exhaustive comparison - including tuning of all different possible parameters for each tool- was out of scope for this study.

Our results will enable the discovery of novel chloroplast genomes as well as an assembly of inter/intra-individual differences in the respective chloroplast genomes.

\section*{Results}
\subsection*{Performance metrics}
%\subsection*{Computational Resources}
%Table and figure of performance evaluation. Including RAM, I/O, runtime, parallelization ability, CPU usage. Free text description of results, range, mean/median, effect of input size, threads parameter, ...

\subsubsection*{Time requirements}
In terms of run time, massive differences between the different tools have been observed. Apart from tool-specific differences, input data and number of threads had huge impact. The observed run times varied from a few minutes to several hours (\cref{fig:performance_runtime}).

Some assemblies failed to finish within the time-limit we set (\SI{48}{\hour}). 
On average the longest time to generate the assemblies was taken by \ioga{} and \fp{} followed by \oa{} and \go{} the most time efficient tool was \ce{}, which on average is a little faster than \np{} and \cassp{}.
%% multithreading

Not all the tools were able to benefit from having access to multiple threads. Both \np{} and \oa{} take about the same time independent of being able to utilize \numlist[list-final-separator={, or }]{1;2;4;8} threads. \cassp{}, \ce{}, \go  and \fp{} all profit from multi-threading (\cref{fig:performance_runtime,fig:performance_memory_cpu,tab:perform25K_suppl,tab:perform250K_suppl,tab:perform2.5M_suppl}).

\subsubsection*{Memory and CPU Usage }
The peak and mean CPU usage, as well as peak memory and disk usage have been recorded for all assemblers based on the same input data set and number of threads to use (\cref{fig:performance_memory_cpu,tab:perform25K_suppl,tab:perform250K_suppl,tab:perform2.5M_suppl}).
Mainly, the size of the input data influenced the peak memory usage with the exception of \ce{} and \ioga{}.
Those two assemblers seems to have a memory usage pattern, which is less influenced by the size of the data.
The number of allowed threads had only a limited impact on the peak memory usage.
Nevertheless, all programs profit by a higher number of threads, if the size of the input data was increased.
In contrast, the disk usage is independent from input size and number of threads for all assemblers.
%but the factor between the programs exceeds a factor of \num{100}.

\subsection*{Qualitative}
The user experience of most tools was evaluated as mainly \good{} (\cref{tab:resultsQual}). However, a few critique points remained.
Two minor dependencies were missing in the \go{} installation instructions and there was no test data available.
Additionally, an issue occurred when running it on a \plant{At} data set.
We are currently in the process of resolving this with the authors.

The \fp{} installation instructions were missing some dependencies.
Like \go{}, \fp{} does not offer a test data set or a tutorial, except for some example commands.

The \oa{} installation instructions did not work.
We found some issues, which are probably related to the requirement of \texttt{Python~3.7}.
There is a tutorial where sample data is available.
However, following the instructions resulted in a segmentation fault.
We found a workaround for this bug and contacted the authors.

The main critique point of \np{} was the lack of a test data set with instructions.
This was fixed by the authors after we contacted them.
Additionally, \np{} uses a custom license, where an OSI approved license would be preferred.

>>>>

\usepackage{siunitx}
\DeclareSIUnit \basepair{bp}
\DeclareSIUnit \byte{B}
\DeclareSIPrefix\mebi{Mi}{10}
\DeclareSIPrefix\gibi{Gi}{10}

@@

% \usepackage{todonotes}
% \setlength{\marginparwidth}{3cm}

% \newcounter{todocounter}
% \newcommand{\ff}[1]
% {\stepcounter{todocounter}
%  \todo[color=blue!40,author=For Frank]{\thetodocounter: #1}
%  }
% \newcommand{\ak}[1]
% {\stepcounter{todocounter}
%  \todo[color=green!40,author=Arthur]{\thetodocounter: #1}
%  }
\newcommand{\todo}[1]{\textcolor{red}{\bfseries(ToDO: #1})}

@@

\usepackage[acronym,hyperfirst=false]{glossaries}
\newacronym{joss}{JOSS}{Journal of Open Source Software}
\newacronym{lsc}{LSC}{Large Single Copy}
\newacronym{ssc}{SSC}{Small Single Copy}
\newacronym{ir}{IR}{Inverted Repeat}
\newacronym{gui}{GUI}{graphical user interface}

\glsdisablehyper
\makeglossaries

@@

%\usepackage[noabbrev]{cleveref}
\usepackage{cleveref}
\crefname{figure}{Figure}{Figures}
\crefname{table}{Table}{Tables}

@@

\author[
    addressref={aff1,aff3},
    %noteref={n1},
    email={jan.freudenthal@uni-wuerzburg.de}
    ]{\inits{JAF}\fnm{Jan A} \snm{Freudenthal}}
\author[
    addressref={aff1,aff2},
    email={simon-pfaff@gmx.de}
    ]{\inits{SP}\fnm{Simon} \snm{Pfaff}}
\author[
   addressref={aff1,aff3},
   email={niklas.terhoeven@uni-wuerzburg.de}
]{\inits{NT}\fnm{Niklas} \snm{Terhoeven}}
\author[
   addressref={aff1},
   email={arthur.korte@uni-wuerzburg.de}
]{\inits{AK}\fnm{Arthur} \snm{Korte}}
\author[
   addressref={aff1,aff3,aff4 },
   noteref={n2},
   email={markus.ankenbrand@uni-wuerzburg.de}
]{\inits{MJA}\fnm{Markus J} \snm{Ankenbrand}}
\author[
   addressref={aff1,aff2,aff5},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   noteref={n2},                        % id's of article notes, if any
   email={frank.foerster@uni-wuerzburg.de}   % email address
]{\inits{FF}\fnm{Frank} \snm{Förster}}

@@

\address[id=aff3]{%
  \orgname{AnaLife Data Science},
  \street{Wiesengrund 16},
  \postcode{97295 Waldbrunn}
  \city{W\"{u}rzburg},
  \cny{Germany}
}
\address[id=aff2]{%                           % unique id
  \orgname{Department of Bioinformatics, University of W\"{u}rzburg}, % university, etc
  \street{Biozentrum, Am Hubland},                     %
  \postcode{97074},                                % post or zip code
  \city{W\"{u}rzburg},                              % city
  \cny{Germany}                                    % country
}
\address[id=aff4]{%
  \orgname{Chair of Cellular and Molecular Imaging, Comprehensive Heart Failure Center, University Hospital W\"{u}rzburg},
  \street{Josef-Schneider-Str.\ 2},
  \postcode{97080}
  \city{W\"{u}rzburg},
  \cny{Germany}
}

\address[id=aff5]{%
  \orgname{Fraunhofer IME-BR},
  \street{Winchester Str.\ 2},
  \postcode{35394}
  \city{Gie\ss{}en},
  \cny{Germany}
}

@@

\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
\note[id=n2]{Corresponding author} % note, connected to author
\end{artnotes}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstractbox}
\begin{abstract} % abstract
%One or two sentences providing a basic introduction to the field, comprehensible to a scientist in any discipline.
Chloroplasts are intracellular organelles that enable plants to conduct photosynthesis. They arose through the symbiotic integration of a prokaryotic cell into an eukaryotic host cell and still contain their own genomes with distinct genomic information.
Plastid genomes accommodate essential genes and are regularly utilized in biotechnology or phylogenetics.
%Two to three sentences of more detailed background, comprehensible to scientists in related disciplines.
Different assemblers that are able to asses the plastid genome, have been developed.
These assemblers often use data of whole genome sequencing experiments, which usually contain reads from the complete chloroplast genome as well, as input.

%One sentence clearly stating the general problem being addressed by this particular study.
The performance of different assembly tools has never been systematically compared. %to each other concerning success rate or computational requirements.
%One sentence summarising the main result (with the words “here we show” or their equivalent).
Here we present a benchmark of seven chloroplast assembly tools, capable to succeed in more than \SI{60}{\percent} of known real data sets.
%Two or three sentences explaining what the main result reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.
Our results show significant differences between the tested assemblers in terms of generating whole chloroplast genome sequences and computational requirements.
The examination of more than \num{100} data sets with unknown plastid genomes led to the assembly of novel chloroplast genomes in almost \SI{50}{\percent} of the cases.
%\todo{on the novel set we only have 49/105 which is roughly 47 percent}
%One or two sentences to put the results into a more general context. 
We created docker images for each tested tool that are freely available for the scientific community and ensure reproducibility of the analyses.
These containers allow to analyze and screen data sets for chloroplast genomes using standard computational infrastructure.
%Two or three sentences to provide a broader perspective, readily comprehensible to a scientist in any discipline, may be included in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced by their inclusion. 
Thus, large scale screening for chloroplasts as hidden treasures within genomic sequencing data is feasible.

@@

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Introduction}
\subsection*{General introduction and motivation}
Chloroplasts are essential organelles present in the cells of plants and autotrophic protists, which enable the conversion of light energy into chemical energy via photosynthesis. They harbor their own prokaryotic type of ribosomes and a circular DNA genome that varies in size between \SIrange{120}{160}{\kilo\basepair} \cite{palmer_1985}.
Because of their small size, chloroplast genomes were one of the first targets for sequencing projects.
The first chloroplast genome sequences were obtained in 1986 \cite{ohyama_chloroplast_1986,shinozaki_complete_1986}.
These early efforts elucidated the general genome organization and structure of the chloroplast DNA and have been reviewed previously \cite{wicke_evolution_2011,green_chloroplast_2011}.
Chloroplast genomes are widely used for evolutionary analyses \cite{martin_plastid_2010,xiao-ming_inferring_2017}, barcoding \cite{kress_use_2005,hollingsworth_dna_2009,de_vere_dna_2015}, and meta-barcoding \cite{bell_review_2016,deiner_environmental_2017}.
Interesting features of chloroplast genomes include their small size (\SIrange{120}{160}{\kilo\basepair},\cite{palmer_1985}), due to endosymbiotic gene transfer \cite{martin_evolutionary_2002,timmis_endosymbiotic_2004}, and the low number of \numrange{100}{120} genes that are encoded within the genome \cite{wicke_evolution_2011}.
Despite the overall high sequence conservation of the chloroplast genome, there are striking differences in the gene content between different autotroph groups, exemplified by the loss of the whole \genename{ndh} gene family in Droseraceae \cite{nevill_plastome-wide_2019}). Even more extreme evolutionary cases, where chloroplasts show a very low GC content and a modified genetic code have been described \cite{su_novel_2019}.

Structurally, two inverted repeats (\glspl{ir}) named \gls{ir}\textsubscript{A} and \gls{ir}\textsubscript{B} of \SIrange{10}{76}{\kilo\basepair} divide the chloroplast genome into a 
%long single copy (\gls{lsc}) and a short single copy (\gls{ssc}) region
\gls{lsc} and a \gls{ssc} region
\cite{palmer_1985}, which complicates genome assembly with short read technologies\cite{Wang2018}.
Moreover, the existence of different chloroplasts within a single individual, and thus multiple different chloroplast genomes, have been described for various plants \cite{corriveau_1988,Chat2002,Scar2016}. This phenomenon - called heteroplasmy - is only poorly understood in terms of its origin and evolutionary importance, but it impacts the assembly of whole chloroplast genomes.

Nonetheless, given its small size, it is still much easier to decipher a complete chloroplast genome than a complete core genome. Consequently, many comparative genomic approaches target the chloroplast genome.
For example the \plant{At} core genome is approximately \SI{125}{\mega\basepair} in length \cite{schmuths2004,cao2011} while the size of the \plant{At} chloroplast genome at \SI{154}{\kilo\basepair} is more than \SI{800} times smaller \cite{sato1999}.

Each single chloroplast contains several hundred copies of its genome \cite{kumar_2014,bendich_1987}.
Therefore, many plant core genome sequencing projects contain reads that originate from chloroplasts as a by-product and permit the assembly of chloroplast genomes.
Such sequences are available from databases such as the Sequence Read Archive at NCBI \cite{sra2010}.

Complete chloroplast genomes can be used as super-barcodes \cite{coissac_barcodes_2016}, both for biotechnology applications and genetic engineering \cite{daniell_chloroplast_2016}.
Furthermore, the availability of whole chloroplast genomes would enable large scale comparative studies \cite{tonti-filippini_what_2017}.

\subsection*{Approaches to extracting chloroplasts sequences from whole genome data}

Different strategies have been developed to assemble chloroplast genomes \cite{twyford_strategies_2017}.
In general, obtaining a chloroplast genome from whole genome sequencing (WGS) data requires two steps: (1) extraction of chloroplast reads from the sequencing data;  
(2) assembly and resolution of the special circular structure including the \glspl{ir}.
The extraction of chloroplast reads can be achieved by mapping the reads to a reference chloroplast. \cite{Vinga2012}.
%\todo{gibt es dazu ein beispiel in unserem set? (niklas)}
A different approach that does not depend on the availability of a reference chloroplast, uses the higher coverage of reads originating from the chloroplast \cite{Chan2013}.
Here, a $k$-mer analysis can be used to extract the most frequent reads.
An example for this is implemented in \ce{} \cite{j_ankenbrand_chloroextractor:_2018}. %\todo{ce ist hier der einzig explizit genannte Assembler, das wirkt wie bias :) entfernen? Nope... Steht klar ein Beispiel und wir haben nun mal den angesehen... Würde ich jetzt einmal so drin lassen. Alternativ nehmen wir noch einen dazu}
%Arthur : Ich würds so lassen.
A third method, which is for example used by \np{} \cite{dierckxsens_novoplasty:_2017}, combines both approaches by using a reference chloroplast as seed and simultaneously assembling the reads based on $k$-mers.

@@

\subsection*{Purpose and scope of this study}
The goal of this study was to compare the effectiveness and efficiency of existing open source command-line tools to perform a de-novo assembly of whole chloroplast genomes from raw genomic data.
We only compared tools that require minimal configuration, which includes no need for extensive data preparation, no need for a specific reference (apart from \plant{At}), no need to change default parameters, and no manual finishing.
We further restricted our benchmark to paired end Illumina data as the sole input, as  these are routinely generated by modern sequencing platforms \cite{Goodwin2016}.

Thus our analyses reflect the most common use cases: (1) trying a tool quickly without digging into options for fine tuning; (2) large scale automatic applications.
We tested all tools on more than \num{100} real data sets for species without published chloroplast genomes.
The performance of most tools might be significantly improved by optimizing parameters for each data set specifically, but this exhaustive comparison - including tuning of all different possible parameters for each tool- was out of the scope of this study.

To summarize, we provide new chloroplast genome sequences for many species and demonstrated the ability to discover and assemble novel chloroplast genomes as well as asses inter/intra-individual differences in the respective chloroplast genomes.

\section*{Results}
\subsection*{Performance metrics}
%\subsection*{Computational Resources}
%Table and figure of performance evaluation. Including RAM, I/O, runtime, parallelization ability, CPU usage. Free text description of results, range, mean/median, effect of input size, threads parameter, ...
All described tools have been tested with regard to their assembly time, memory and CPU utilization.

\subsubsection*{Time requirements}
Massive differences between the different tools were observed in terms of the run time for the assembly. Apart from tool-specific differences, input data and number of threads used had a huge impact on the time requirement. The observed run times varied from a few minutes to several hours (\cref{fig:performance_runtime}).

Some assemblies failed to finish within our time limit of \SI{48}{\hour}. 
On average, the longest time to generate an assembly was taken by \ioga{} and \fp{} followed by \oa{} and \go{}.
The most time efficient tool was \ce{}, which was a little faster than \np{} and \cassp{}.
%% multithreading

Not all tested tools were able to benefit from having access to multiple threads.
Both \np{} and \oa{} required almost the same time independent of being allowed to utilize \numlist[list-final-separator={, or }]{1;2;4;8} threads.
In contrast, \cassp{}, \ce{}, \go  and \fp{} all profited from multi-threading settings  (\cref{fig:performance_runtime,fig:performance_memory_cpu,tab:perform25K_suppl,tab:perform250K_suppl,tab:perform2.5M_suppl}).

\subsubsection*{Memory and CPU Usage }
The peak and mean CPU usage, as well as peak memory and disk usage were recorded for all assemblers based on the same input data set and number of threads  (\cref{fig:performance_memory_cpu,tab:perform25K_suppl,tab:perform250K_suppl,tab:perform2.5M_suppl}).
In general, the size of the input data influenced the peak memory usage with the exception of \ce{} and \ioga{}.
Those two assemblers showed a memory usage pattern, which was less influenced by the size of the data.
The number of allowed threads had only a limited impact on the peak memory usage.
All programs profited from a higher number of threads, if the size of the input data was increased concerning their memory and CPU usage footprint.
%\todo{oben steht nicht alle? (niklas)}
% Frank: Hier geht es um Memory und CPU-Usage... Oben nur um Runtime
In contrast, the disk usage was independent of the size of the input data and the number of threads for all assemblers.

%but the factor between the programs exceeds a factor of \num{100}.

\subsection*{Qualitative}

On average, the user experience in terms of installation and running the analyses was evaluated as \good{} for all tools (\cref{tab:resultsQual}).

However, we discovered the following slight problems:

Two minor dependencies were missing in the \go{} installation instructions and there were no test data available \cite{go_issue_10}.
Additionally, an issue occurred when running it on one particular \plant{At} data set.
This was resolved after contact with the authors via GitHub \cite{go_issue_11}.
%
